Modeling browser-based distributed evolutionary computation systems

### 2015-02-06 modeling-volunteer-computing-gecco.tex
Here goes a compiling version

\begin{abstract}
From the era of big science we are back to the "do it yourself", where
you do not have any money to buy clusters or subscribe to grids but
still have algorithms that crave many computing nodes and need them to
measure scalability. Fortunately, this coincides with the era of big
data, cloud computing, and browsers that include JavaScript virtual
machines. Those are the reasons why this talk will focus on two
different aspects of volunteer or freeriding computing: first, the
pragmatic: where to find those resources, which ones can be used, what
kind of support you have to give them; and then, the theoretical: how
evolutionary algorithms can be adapted to a environment in which nodes
come and go, have different computing capabilities and operate in
complete asynchrony of each other. 
\end{abstract}

@@

Most theory in parallel computing has been devoted to predict and optimize the performance in systems where the number of nodes, their connections, and the time every one is devoting to the computation is known in advance. However, even if Big Science is not over, the era of Citizen science has started (with SETI@home and then BOINC) and it offers a vast amount of computational resources to attract, if only you know how. But there is a challenge: knowing, or at least having a ballpark, of how your algorithm is going to perform in this uncertain environment, where none of the factors is known: neither the number of nodes, through how they are connected, to how long are they going to be focused on doing what you want them to.

@@

>>>>

\begin{abstract}
From the era of big science we are back to the "do it yourself", where
you do not have any money to buy clusters or subscribe to grids but
still have algorithms that crave many computing nodes and need them to
measure scalability. Fortunately, this coincides with the era of big
data, cloud computing, and browsers that include JavaScript virtual
machines. Those are the reasons why this paper will focus on two
different aspects of volunteer or freeriding computing: first, the
pragmatic: where to find those resources, which ones can be used, what
kind of support you have to give them; and then, the theoretical: how
evolutionary algorithms can be adapted to a environment in which nodes
come and go, have different computing capabilities and operate in
complete asynchrony of each other. We will examine the setup needed to
create a very simple distributed evolutionary algorithm using
JavaScript and then find a model of how users react to it by
collecting data from several experiments featuring different classical
benchmark functions. 
\end{abstract}

@@

Most theory in parallel computing has been devoted to predict and
optimize the performance in systems where the number of nodes, their
connections, and the time every one is devoting to the computation is
known in advance. However, even if Big Science is not really over and
it is slated to come back, the era of
Citizen Science already started a few years ago (with SETI@home \cite{david-seti:home} and then BOINC \cite{boinc_grid04}) and it
offers a vast amount of computational resources to attract, if only
you know how. But there is a challenge: knowing, or at least having a
ballpark, of how your algorithm is going to perform in this uncertain
environment, where none of the factors is known: neither the number of
nodes, through how they are connected, to how long are they going to
be focused on doing what you want them to. That is why some effort is
needed to first understand the dynamics of the decision to participate
in an experiment that requires you to click on a link and then stay
for a while looking at the screen (or just leave it there running).

@@

\begin{figure*}[h!tb]
  \centering
  \includegraphics[trim=0cm 2cm 0cm 3cm, width=10cm]{img/screenshot.png}
  \label{fig:sshot}
  \caption{Screenshot of the talk that includes at the graph
    that shows the progression of the evolutionary
    algorithm at the top right corner}
\end{figure*}
